---
title: "Milestone 3"
author: "Adelson Aguasvivas"
date: "February 26th, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(sf)
library(rgdal)

```

# About Draft

#### [Access my Github repository here!](https://github.com/aaguasvivas/MS_3)

For my final project, I want to focus on the history and effects of racism and
discrimination, specifically through redlining. I will look at how redlining
affected the geography of New York City, where I am from. I will be focusing on
the five different boroughs (the Bronx, Queens, Manhattan, Brooklyn and Staten
Island) and see how the city has been divided. I will use shapefile data from
1937-1940, to look at these divisions, which use different rankings to
categorize geographical locations. These geographical locations were associated
with race, with some of the "hazardous" zones often being those with the highest
Black and Hispanic populations. I will be using census data to determine how
those zones are still affected today, and what the change has been throughout
the last 80 years. My expectation is that the zones that were redlined because
of people of color will be some of the poorest places still today. I expect a
lot of these areas to still have a high concentration of people of color. If
not, I will look to see whether this change from being a city that was redlined,
to now being more wealthy, is one of the results of gentrification. Right now, I
have collected a lot of the map data through shapefiles, which I am not sure how
to work with yet, and I am actively trying to learn about. I am using census
data which we have worked with in class before. I am looking for other sorts of
data that will help me prove my claim or give me some sort of guidance as to
what kinds of observations I can make.

Right now, my plan is to examine this data and talk to my TF, Mitchell, to see
what kinds of comparisons are useful when working with shapefile maps and census
data. Something I am worried about and will keep working through is whether I am
reading and processing this data correctly and whether I will be able to draw
useful observations moving forward. As of right now, I will be using a lot of
what I have mentioned here as part of my future About Page, as well as hopefully
a lot of the data gathered into my raw-data subdirectory. I will be taking the 
datacamp course that focuses on using sf, as well as a few other courses, 
maybe around mapping.

# Datasets Used & Links

I used the following datasets:

- Mapping Inequality Redlining in New Deal America Shapefiles. 
[(here)](https://dsl.richmond.edu/panorama/redlining/#loc=6/37.484/-79.955&text=downloads)
    
- New York City Census Data: Demographic, Economic, and Location Data for 
Census Tracts in NYC. 
[(here)](https://www.kaggle.com/muonneutrino/new-york-city-census-data/data)

```{r data-processing, include=FALSE}

# Here I will load in and process the maps, as well as the census data.
# First I start off my processing NYC location data.

NY_census_block_loc <- 
  read_csv("raw-data/new-york-city-census-data/census_block_loc.csv", 
           col_names = TRUE) %>%
  clean_names() %>%
  filter(county == "Queens" | county == "Bronx" | county == "New York")

# Here, I process a more traditional census dataset/table with lots of
# information

NY_census_tracts <-
  read_csv("raw-data/new-york-city-census-data/nyc_census_tracts.csv", 
           col_names = TRUE) %>%
  clean_names()

# Below I create a variable where I read in the shapefiles of each of the five
# boroughs, then I use this variable along with ggplot to graph a map of the
# boroughs. The next step would be to creative a comprehensive visual of housing
# segregation in 1937-1940 versus what the low income parts are now using the
# census data.

NY_Bronx_1938 <- 
  st_read("raw-data/NYBronx1938/cartodb-query.shp")

Bronx_plot <-
  ggplot() + 
  geom_sf(data = NY_Bronx_1938) + 
  ggtitle("Bronx") + 
  coord_sf()

NY_Brooklyn_1938 <- 
  st_read("raw-data/NYBrooklyn1938/cartodb-query.shp")

Brooklyn_plot <-
  ggplot() + 
  geom_sf(data = NY_Brooklyn_1938) + 
  ggtitle("Brooklyn") + 
  coord_sf()

NY_Manhattan_1937 <- 
  st_read("raw-data/NYManhattan1937/cartodb-query.shp")

Manhattan_plot <-
  ggplot() + 
  geom_sf(data = NY_Manhattan_1937) + 
  ggtitle("Manhattan") + 
  coord_sf()

NY_Queens_1938 <- 
  st_read("raw-data/NYQueens1938/cartodb-query.shp")

Queens_plot <-
  ggplot() + 
  geom_sf(data = NY_Queens_1938) + 
  ggtitle("Queens") + 
  coord_sf()

NY_StatenIsland_1940 <- 
  st_read("raw-data/NYStatenIsland1940/cartodb-query.shp")

StatenIsland_plot <-
  ggplot() + 
  geom_sf(data = NY_Manhattan_1937) + 
  ggtitle("Staten Island") + 
  coord_sf()

```

