---
title: "Final"
author: "Adelson Aguasvivas"
date: "April 23rd, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)


```

# About Draft

#### [Access my Github repository here!](https://github.com/aaguasvivas/MS_4)

For my final project, I want to focus on the history and effects of racism and
discrimination, specifically through redlining. I will look at how redlining
affected the geography of New York City, where I am from. I will be focusing on
the five different boroughs (the Bronx, Queens, Manhattan, Brooklyn and Staten
Island) and see how the city has been divided. I will use shapefile data from
1937-1940, to look at these divisions, which use different rankings to
categorize geographical locations. These geographical locations were associated
with race, with some of the "hazardous" zones often being those with the highest
Black and Hispanic populations. I will be using census data to determine how
those zones are still affected today, and what the change has been throughout
the last 80 years. My expectation is that the zones that were redlined because
of people of color will be some of the poorest places still today. I expect a
lot of these areas to still have a high concentration of people of color. If
not, I will look to see whether this change from being a city that was redlined,
to now being more wealthy, is one of the results of gentrification. Right now, I
have collected a lot of the map data through shapefiles, which I am not sure how
to work with yet, and I am actively trying to learn about. I am using census
data which we have worked with in class before. I am looking for other sorts of
data that will help me prove my claim or give me some sort of guidance as to
what kinds of observations I can make.

Right now, my plan is to examine this data and talk to my TF, Mitchell, to see
what kinds of comparisons are useful when working with shapefile maps and census
data. Something I am worried about and will keep working through is whether I am
reading and processing this data correctly and whether I will be able to draw
useful observations moving forward. As of right now, I will be using a lot of
what I have mentioned here as part of my future About Page, as well as hopefully
a lot of the data gathered into my raw-data subdirectory. I will be taking the 
datacamp course that focuses on using sf, as well as a few other courses, 
maybe around mapping.

# Datasets Used & Links

I used the following datasets:

- Mapping Inequality Redlining in New Deal America Shapefiles. 
[(here)](https://dsl.richmond.edu/panorama/redlining/#loc=6/37.484/-79.955&text=downloads)
  - This dataset provides us with accurate visuals of what readlining was like
  from 1937-1940, in the different parts of New York. We have access to several
  different parts of New York, but I have decided to focus on the five boroughs,
  not only because these are the places I am more familiar with being from New
  York, but also because this is where some of the worst redlining in the
  country happened about 80 years ago. I hope to explore this data and compare
  it to recent census data, to see which patterns and observations are relevant
  and important to discuss.
    
- New York City Census Data: Demographic, Economic, and Location Data for 
Census Tracts in NYC. 
[(here)](https://www.kaggle.com/muonneutrino/new-york-city-census-data/data)
  - This is the most recent census data I could find which provides huge amounts
  of information, such as where the people live, income levels, education, rate
  of poverty, income per capita, race, gender, among other things. I will use
  this information and compare it to the redlining maps from 80 years ago to see
  ouch patterns have developed because of redlining and how have people been
  affected by it even 80 years later.
  
- Tidycensus Library
  - I used some of the techniques we have learned in class, like get_acs, to
  plot the median income for the five boroughs in New York city in 2018, as well
  as each of their racial geographies.
  
  
  
```{r graph-side-by-side, include=FALSE}

# For the milestone I wanted to create another graph where I compare the change
# in the racial demographics of the five boroughs using webscrapped data from
# wikipedia, so that we can compare the different racial distributions and how
# this plays a part in median income. I started by adding options to have the
# numbers not be in scientific notation. To scrape the data from wikipedia, I
# saved the url, read it in, read the html, found the xpath, made a table then
# cleaned the names using the janitor package. Then, I began to organize my data
# to plot it, first, by selecting the relevant variables, pivoting to organize
# the data better, transforming the population column to numeric, making all the
# NAs to zeros, so that we do not lose this data, even though the census didn't
# keep track of some specific populations like Hispanic/Latinxs at one point.
# Then, I used ggplot, to set the x axis to the year and the y axis to the
# population, using geom_point to create a scatterplot, so we can visualize
# change over time, setting the dots to be the race, and using position_jitter,
# to avoid overplotting, as well as adding alpha, so that we can see through the
# points better, so that we do not miss any of them. I used labs, to give the
# graph a meaningful title, subtitle, and axis names. I used
# scale_color_discrete, to fix the legend and make it display better
# information. Finally, I used scale_x_continuous and scale_y_continuous to set
# the breaks and labels for the axis.

options(scipen=5)

website = 'https://en.wikipedia.org/wiki/Demographic_history_of_New_York_City#The_Bronx'

table_bronx <- website %>%
  read_html() %>%
  html_node(xpath = '//*[@id="mw-content-text"]/div/table[2]') %>%
  html_table(fill = TRUE) %>%
  clean_names()
  
change_over_time_bronx <- table_bronx %>%
  select(year, white_includes_white_hispanics, 
         black, asian, hispanic_latino, foreignborn) %>%
  pivot_longer(-year, names_to = "race", values_to = "population") %>%
  transform(population = as.numeric(gsub(",", "", population))) %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>%
  ggplot(aes(x = year, y = population, color = race)) +
  geom_line() +
  geom_point(position = position_jitter(width = 0.3, 
                                        height = 0), 
             alpha = 0.5) +
  labs(title = "Change in Demographics Over Time in the Bronx, New York",
       subtitle = "Data Webscrapped From Wikipedia. From 1910-2010.",
       x = "Year",
       y = "Population") +
  theme_classic() +
  scale_color_discrete(name = "Race/Census Categories",
                            labels = c("Asian", "Black", "Foreign Born", 
                                       "Hispanic or Latino", "White")) +
  scale_x_continuous(breaks = c(1910, 1920, 1930, 1940, 1950, 
                                1960, 1970, 1980, 1990, 2000, 2010)) +
  scale_y_continuous(breaks = c(0, 250000, 500000, 750000, 1000000, 1250000),
                     labels = c("0", "250,000", "500,000", "750,000",
                                "1,000,000", "1,250,000"))

```
```{r display-change, echo=FALSE}

# Displays graph on knitted html 

change_over_time_bronx

```


# Progress Update Since Milestone 5:

**LINK TO THE REPO: https://github.com/aaguasvivas/MS_6**

**LINK TO THE SHINY APP: https://adelson-aguasvivas.shinyapps.io/ms_6_change_over_time_bronx_animated/**

```{r ms-6, include=FALSE}

animated_change <- change_over_time_bronx +
  transition_reveal(year) +
  labs(title = "Change in Demographics Over Time in the Bronx. Year: {as.integer(frame_along)}")

# anim_save("bronx_change_animation.gif", path = "ms_6/", 
#          animation = animated_change)

nycounties <- rgdal::readOGR("raw-data/NYManhattan1937/geojson/NYManhattan1937.geojson")

# Look at levels of Home Owners’ Loan Corporation (HOLC) “Residential Security”
# grades (D is red)

nycounties$holc_grade

## Create color factor palette which makes `A' green, 'B' blue, 'C' yellow

factpal <- colorFactor(c("green","blue","yellow","red"), nycounties$holc_grade)

# Map redlining in Manhattan
# Plots holc_grade using factpal palette

manhattan_redlining1937 <- leaflet(nycounties) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
    fillColor = ~factpal(holc_grade))

manhattan_redlining1937

```
```{r ny1950census, include=FALSE}

# Import Census Tract Data as data.frame Consult
# nhgis0013_ds82_1950_tract_codebook.txt codebook file to determine that column
# B0F001 contains Median Income in 1949. Rename this MedianHHIncome for clarity.
# Will need to figure out what the income variables are called for each new year
# of data you download from the codebook that comes with each NHGIS download.
# Select only MedianHHIncome and GISJOIN variables. GISJOIN is the variable that
# allows you to join the census tract median income data with the shapefile
# data. Also drop census tracts which reported 0 median income

data1950 <- read.csv("raw-data//nhgis0013_csv/nhgis0013_ds82_1950_tract.csv")%>%
  filter(STATE == "New York" & B0F001 > 0)%>%
  mutate(MedianHHIncome1949 = B0F001)%>%
  select(GISJOIN,MedianHHIncome1949)

# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
# layer is the filename of the .shp file inside the
# folder dsn points to. 

tracts1950 <- sf::st_read(dsn = "raw-data/nhgis0012_shape/nhgis0012_shapefile_tl2008_us_tract_1950",
                         layer = "US_tract_1950_conflated")

# *Select New York observations only using NHGISST variable NHGISST should be
# consistent between census years, but you might need to double-check this if
# you try 1960, 1970. *The NHGISST code is 360 for New York because New York's
# FIPS code is 360. Google New York FIPS code for more information about the
# FIPS system. *Need to have spdplyr package loaded to use tidyverse commands on
# SpatialPolygonsDataFrame aka `filter' and 'mutate' *Join the data1950 data
# with the MedianHHIncome variable by 'GISJOIN' so that the tract lines and
# income data are in one object *Subset to Manhattan county, which has FIPS code
# 061 *Can see other NY county fips codes at
# https://simple.wikipedia.org/wiki/List_of_counties_in_New_York

tracts1950 <-
  tracts1950 %>% filter(NHGISST == "360" & COUNTY=="061")%>%
  merge(data1950, "GISJOIN")


# Set projection of tracts dataset to `projection` required by leaflet

tracts1950<-sf::st_transform(tracts1950,  crs="+init=epsg:4326")

# Condense size of data for faster processing

tracts1950<- rmapshaper::ms_simplify(tracts1950)

# Set palette color

pal <- colorNumeric("viridis", NULL)

#  Plot the data

leaflet(tracts1950) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
              fillColor = ~pal(MedianHHIncome1949)) %>%
  addLegend(pal = pal, values = ~MedianHHIncome1949, opacity = 1.0)



```